# Reproducible Research: Course Project 1

## Loading and preprocessing the data
```{r loaddata}
unzip(zipfile="activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="Total number of steps taken per day")
```

Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("Average number of steps taken")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
averages[which.max(averages$steps),]
```

## Imputing missing values

There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r how_many_missing}
missing <- is.na(data$steps)
# the number of missing values
table(missing)
```

All of the missing values are filled in with mean value for that 5-minute
interval.

```{r}
# Replace each missing value with the mean value of its 5-minute interval
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```
Now, using the filled data set, let's make a histogram of the total number of steps taken each day and calculate the mean and median total number of steps.

```{r}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="Total number of steps taken each day")
mean(total.steps)
median(total.steps)
```

Mean and median values are higher after imputing missing data. The reason is
that in the original data, there are some days with `steps` values `NA` for 
any `interval`. The total number of steps taken in such days are set to 0s by
default. However, after replacing missing `steps` values with the mean `steps`
of associated `interval` value, these 0 values are removed from the histogram
of total number of steps taken each day.

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” 
indicating whether a given date is a weekday or weekend day.

```{r}
filled.data$dateType <-  ifelse(as.POSIXlt(filled.data$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
averagedfilled.data <- aggregate(steps ~ interval + dateType, data=filled.data, mean)
ggplot(averagedfilled.data, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(dateType ~ .) +
  xlab("5-minute interval") + 
  ylab("Average number of steps")
```
